Uncertainty on the galaxy-halo connection for Lyman-$\alpha$ emitters at $z=3.1$

### 2016-06-14 paper/ms_old.tex
previous version + diff

\begin{abstract}
 In this paper we study the impact of cosmic variance and observational uncertainties
in constraining the mass and occupation fraction, \focc, of
dark matter halos hosting \ly\ Emitting  Galaxies (LAEs) at high redshift. 
To this end, we construct mock catalogs from an N-body simulation to match the 
typical size of observed fields at $z=3.1$ ($\sim 1 {\rm deg^2}$).
In our model a dark matter halo with mass in the range $\mmin
<M_{\mathrm h}<\mmax$ can only host one detectable LAE at most.    
We proceed to  explore the parameter space determined by \mmin, \mmax\ and \focc\
with a Markov Chain Monte-Carlo algorithm using the angular correlation function (ACF) and the LAEs number density  as observational constraints. 
We find that the preferred minimum and maximum masses in our model span
a wide range $10^{10.0}\hMsun\leq \mmin \leq 10^{11.1}\hMsun$ ,
$10^{11.0}\hMsun\leq \mmax \leq 10^{13.0}\hMsun$; followed by a
wide range in the occupation fraction $0.02\leq \focc \leq 0.30$.   
As a consequence the median mass, $M_{50}$, of all the consistent models has a large uncertainty $M_{50} =
3.16^{+9.34}_{-2.37}\times 10^{10}$\hMsun.
However, we find that the same individual models have a relatively tight
$1\sigma$  scatter around the median mass $\Delta M_{1\sigma} = 0.55^{+0.11}_{-0.31}$ dex.
We are also able to show that \focc\ is uniquely determined by $M_{\rm
  min}$, regardless of $M_{\rm max}$. 
We argue that upcoming large surveys covering at least $25$ deg$^{2}$
should be able to put tighter constraints on \mmin\ and \focc\ through the LAE number density
distribution width, $W_{1\sigma}$, constructed over several fields of
$\sim 1$ deg$^{2}$.
These results suggest that taking explicitly into account the effect of cosmic
variance using cosmological simulations is a valuable tool to 
interpret LAEs observations at high redshift.
\end{abstract}

@@

A different approach to infer the typical mass of halos hosting
LAEs is based on the spatial clustering information. 
This approach uses the fact
that in CDM cosmologies the spatial clustering of galaxies on large
scales is entirely dictated by the halo distribution
\citep{Colberg00}, which in turn has a strong dependence on halo
mass. 
Using measurements of the angular correlation function of LAEs,
observers have put constraints on the typical mass and occupation
fraction of the putative halos hosting these galaxies
\citep{Hayashino2004,Gawiser07,Nilsson2007,Ouchi2010,Bielby16}. 
In these studies the observations are done on fields of $\sim 1$ deg$^{2}$ and
the conclusions derived on the halo host mass do not elaborate deeply on the
uncertainty resulting from the cosmic variance on these fields. {\bf  \citet{Guaita2010} attempted to analytically 
estimate the uncertainty from cosmic variance 
following \citet{Somerville2004} and \citet{Peebles1980}. Their prescription assumes 
that the correlation function can represented by a power-law $\psi\left(r\right)=\left(r_{0}/r\right)^{\gamma}$. Their estimation of cosmic variance strongly depends on $r_{0}$ and $\gamma$. However, $\gamma$ is only  poorly constrained with current observations which in turns indicates that their cosmic variance estimation could not be precise.}

In this paper we investigate the impact of cosmic variance in
constraining the mass and occupation fraction of halos hosting LAEs at $z=3$.
We build mock surveys from a cosmological N-body simulation to compare them
against the observations of \cite{Bielby16} using the angular
correlation function.  
We use a simple model to populate a halo in the simulation with a LAE   
assuming a minimum, \mmin, and maximum mass, \mmax, for the dark
matter halos hosting LAEs.  We do not  assume  an underlying relation 
between the \ly\  luminosity and the dark matter halo mass.  
This approach bypasses all the physical uncertainties associated to
star formation and radiative transfer. 
We then use the Markov Chain Monte Carlo technique to obtain the
likelihood of the parameters given the observational constraints. {\bf This approach 
allow us to estimate cosmic variance directly from simulations without making any assumption 
on the correlation function behavior.}

@@

\subsection{Observational constraints}
\label{subsec:obs}
\citet{Bielby16} used narrow band imaging to detect 643 LAE candidates
at $z\sim 3$  with Ly-$\alpha$ {\bf rest-frame} equivalent widths $\gtrsim$65\AA\ and a  Ly-$\alpha$ flux limit
of $2\times10^{17}{\rm erg/cm^2/s}$ ($L\sim 7\times10^{42}{\rm erg/s} $). 
Using follow-up spectroscopy they found a 22\% contamination fraction $f_{\rm c}$.
Their observations cover 5 (out of 9) independent and adjacent
fields of the VLT LBG Redshift Survey (VLRS).  
The  total observed  area corresponds to 1.07$\rm deg^2$ that translates to
$\sim$80$^2\hMpc^2$ in a comoving scale. 
\citet{Bielby16} used the NB497  narrow-band filter whose 77\AA\ FWHM
and 154\AA\ Full width  tenth maximum (FWTM) correspond to a total observational comoving depth of
44\hMpc\  and 82\hMpc, respectively.

@@

We use results from the BolshoiP simulation \citep{Bolshoi,BolshoiP} 
performed in a cubic  comoving volume of 250\hMpc\  on a side. 
The dark matter distribution was sampled using  $2048^{3}$
particles.

We chose this simulation based on three requirements. 
First, the simulation should have cosmological parameters consistent
with the most recent constraints from Cosmic Microwave Background
experiments and provide public acces to the data.
Second, the halo mass function from the simulation should be robust
down to masses that include the range of possibilities from previous
observational efforts \cite[e.g. results from][impose a lower bound of
  $10^{10}$\hMsun.]{Hayashino2004,Gawiser2007,Ouchi2010,Bielby16} .
Third, the total volume of the simulation should be at least 10 times
larger the equivalent volume of a single LAE observation field.
These conditions are met as we describe below.

The cosmological parameters are consistent with Planck
results \citep{Planck2014} with a matter density 
$\Omega_{\rm m} = 0.307$, cosmological constant
$\Omega_{\Lambda}=0.693$, dimensionless Hubble constant $h=0.678$, slope
of the power spectrum  $n=0.96$ and normalization of the power
spectrum $\sigma_{8}=0.823$.  
This translates into a particle mass of  $m_{\rm p}=1.5\times 10^{8}$
$h^{-1}$ M$_{\odot}$. 
Data is available to the
public through an online
interface \footnote{\url{http://www.multidark.org/MultiDark/}}
\citep{MultiDark}.

We use halo catalogs constructed with a Bound-Density-Maxima (BDM)
algorithm. 
For each  halo in the box we extract its comoving position and mass.  
We focus our work on halos more massive than $6.0\times
10^{9}$\hMsun\ resolved with at least  $40$ particles 
to guarantee statistical significance and a well behaved halo mass function. 
We do not take into account sub-halos.

@@

In what follows we will describe by the letter ${\mathcal M}$  a model
defined by a particular choice of the two parameters $M_{\rm min}$ and
$M_{\rm  max}$. For each model  ${\mathcal M}$ we define
$\tilde{f}_{\rm occ}$ as the median occupation fraction within the the
mock fields.

@@

The preferred $1-\sigma$ range for the masses is $10.0<\log \mmin<11.2$
and $11.0<\log\mmax<13.0$.
\focc\ is completely determined by \mmin\ from \focc$=$0.011 when
$\log\mmin=10.0$ to $\focc=0.39$ when $\log\mmin=11.15$. 
We compute the power-law dependence between \focc\ and \mmin\ to be 
\begin{equation}
\focc = 0.08\left(\frac{\mmin}{10^{10.3}\hMsun}\right)^{0.77}.
\label{eq:focc}
\end{equation}
We remind the reader that models with $\log \focc>0.00$ ($\focc>1$)
correspond to cases where the halo number density is smaller 
than the number of observed LAEs but are  still considered consistent
because of the expected uncertainty in the median number density of
LAEs due to cosmic variance (see Fig. \ref{fig:cosmicv0} and
\S\ref{subsec:explore}).

{\bf The result in Eq.(\ref{eq:focc}) can be qualitatively understood as
follows.   
A choice of $\focc<1$ means that the halo mass function is scaled down.
If \mmin\ decreases the total number of halos increases because less
massive halos are more abundant.
This requires smaller values of $\focc$ to keep the total
number of halos hosting LAEs equal to the number of LAEs in observations.}

@@

\subsection{Hints of larger uncertainties}

Fig. \ref{fig:like} shows an interesting feature for \mmin. 
The histogram (upper left) showing the probability for this parameter
given the observational constraints does not considerably decreases for lower
masses. 
This raises the following question. Down to which values of
\mmin\ does the probability significantly decreases?

Lower values of \mmin\ would imply a lower \focc\ and larger
uncertainties on the parameters \mmin\, \focc and $M_{50}$.
In other words, the limitations in our priors for \mmin\ (which
set the limitations for the kind of numerical simulation we use) might
be underestimating the uncertainties and slightly biasing our results
for \mmin\, \focc\ and $M_{50}$.

Although Fig. \ref{fig:mmed} suggest that models with lower values of
$M_{\rm 50}$, i.e. lower values of \mmin, have a higher
$\chi^2_{\mathcal{M}}$ and could be thus discarded, we suggest that a
proper test of this hypothesis requires performing new cosmological
simulations to be able to probe \mmin\ masses below our limit of
$6\times  10^{9}$\hMsun.

This highlights the main thesis of this paper. 
Namely, the high impact that cosmic variance can have on constraining
the paremeters of our model. 
With the current limitations in the cosmological simulations available
to the public we cannot significantly decrease or prior on \mmin.
We consider such test beyond the scope of this paper.

This extension to larger priors cannot change other central results of
our paper, such as the relationship between \focc\ and \mmin\, the
narrow mass range for individual models and the possibility to
constraint the halo mass with cosmic variance on the number density
counts. 
However, we suggest that future work building upon the
technique presented in this paper should count with simulations that
extend at least one order of magnitude below our limit $6\times
10^{9}$\hMsun.

@@

This work also highlights the need to explore a wider range on the
prior of \mmin\ if one wants to better estimate the uncertainty on
\mmin, \focc\ and $M_{50}$.  
In this paper we use priors given by previous observations, but our
results suggest that once cosmic variance is explicitly taken into
account, even lower values of \mmin\ might be in fact consistent with
observations. 
We recommend that future work making use of the technique presented
in this paper should extend the prior for \mmin\ to masses at least
one order of magnitude below $6\times 10^{9}$\hMsun.

@@

The authors acknowldge the referee, Nelson Padilla, for useful
remarks and questions that helped us to improve the quality and
clarity of our paper.

@@

\bibitem[{{Guaita} {et~al.}(2010){Guaita}, {Gawiser}, {Padilla},
  {Francke}, {Bond}, {Gronwall}, {Ciardullo}, {Feldmeier}, {Sinawa}, {Blanc},
  \& {Virani}}]{Guaita2010}
{Guaita}, L., {Gawiser}, E., {Padilla}, N., {Francke}, H.,
	{Bond}, N.~A., {Gronwall}, C., {Ciardullo}, R., {Feldmeier}, J.~J., 
	{Sinawa}, S., {Blanc}, G.~A., \& {Virani}, S. 2010, \apj, 714, 255

@@

\bibitem[{{Somerville} {et~al.}(2004){Somerville}, {Lee}, {Ferguson}, {Gardner},
  {Moustakas}, \& {Giavalisco}}]{Somerville2004}
{Somerville}, R.~S., {Lee}, K., {Ferguson}, H.~C.,
	{Gardner}, J.~P., {Moustakas}, L.~A., \& {Giavalisco}, M. 204, \apj,
  600, 171

>>>>

\begin{abstract}
We study the impact of cosmic variance and observational uncertainties
in constraining the mass and occupation fraction, \focc, of
dark matter halos hosting \ly\ Emitting  Galaxies (LAEs) at high redshift. 
We construct mock catalogs from an N-body simulation to match the 
typical size of observed fields at $z=3.1$ ($\sim 1 {\rm deg^2}$).
In our model a dark matter halo with mass in the range $\mmin
<M_{\mathrm h}<\mmax$ can only host one detectable LAE at most.    
We explore the parameter space determined by \mmin, \mmax\ and \focc\
with a Markov Chain Monte-Carlo algorithm using as observational
constraints the angular correlation function (ACF) and the LAEs number density. 
We find that the preferred minimum and maximum masses in our model span
a wide range $10^{10.0}\hMsun\leq \mmin \leq 10^{11.1}\hMsun$ ,
$10^{11.0}\hMsun\leq \mmax \leq 10^{13.0}\hMsun$; followed by a
wide range in the occupation fraction $0.02\leq \focc \leq 0.30$.   
As a consequence the median mass, $M_{50}$, of all models consistent with
observations has a large uncertainty $M_{50} =
3.16^{+9.34}_{-2.37}\times 10^{10}$\hMsun.
However, we find that the same individual models have a relatively tight
$1\sigma$  scatter around the median mass $\Delta M_{1\sigma} = 0.55^{+0.11}_{-0.31}$ dex.
We are also able show that \focc\ is uniquely determined by $M_{\rm
  min}$, regardless of $M_{\rm max}$. 
We argue that upcoming large surveys covering at least $25$ deg$^{2}$
should be able to put tighter constraints on \mmin\ and \focc\ through the LAE number density
distribution width, $W_{1\sigma}$, constructed over several fields of
$\sim 1$ deg$^{2}$.
\end{abstract}

@@

A different approach to infer the typical mass of halos hosting
LAEs is based on the spatial clustering information. 
This approach uses the fact
that in CDM cosmologies the spatial clustering of galaxies on large
scales is entirely dictated by the halo distribution
\citep{Colberg00}, which in turn has a strong dependence on halo
mass. 
Using measurements of the angular correlation function of LAEs,
observers have put constraints on the typical mass and occupation
fraction of the putative halos hosting these galaxies
\citep{Hayashino2004,Gawiser07,Nilsson2007,Ouchi2010,Bielby16}. 
In these studies the observations are done on fields of $\sim 1$ deg$^{2}$ and
the conclusions derived on the halo host mass do not elaborate on the
uncertainty resulting from the cosmic variance on these fields.

In this paper we investigate the impact of cosmic variance in
constraining the mass and occupation fraction of halos hosting LAEs at $z=3$.
We build mock surveys from a cosmological N-body simulation to compare them
against the observations of \cite{Bielby16} using the angular
correlation function.  
We use a simple model to populate a halo in the simulation with a LAE   
assuming a minimum, \mmin, and maximum mass, \mmax, for the dark
matter halos hosting LAEs. We do not  assume  an underlying relation 
between the \ly\  luminosity and the dark matter halo mass.  
This approach bypasses all the physical uncertainties associated to
star formation and radiative transfer. 
We then use the Markov Chain Monte Carlo technique to obtain the
likelihood of the parameters given the observational constraints.

@@

\subsection{Observational constraints}
\label{subsec:obs}
\citet{Bielby16} used narrow band imaging to detect 643 LAE candidates
at $z\sim 3$  with Ly-$\alpha$ equivalent widths of $\gtrsim$65\AA\ and a Ly-$\alpha$ flux limit
of $2\times10^{17}{\rm erg/cm^2/s}$ ($L\sim 7\times10^{42}{\rm erg/s} $). 
Using follow-up spectroscopy they found a 22\% contamination fraction $f_{\rm c}$.
Their observations cover 5 (out of 9) independent and adjacent
fields of the VLT LBG Redshift Survey (VLRS).  
The  total observed  area corresponds to 1.07$\rm deg^2$ that translates to
$\sim$80$^2\hMpc^2$ in a comoving scale. 
\citet{Bielby16} used the NB497  narrow-band filter whose 77\AA\ FWHM
and 154\AA\ Full width  tenth maximum (FWTM) correspond to a total observational comoving depth of
44\hMpc\  and 82\hMpc, respectively.

@@

We use results from the BolshoiP simulation \citep{Bolshoi,BolshoiP} 
performed in a cubic  comoving volume of 250\hMpc\  on a side. 
The dark matter distribution was sampled using  $2048^{3}$
particles. 
The cosmological parameters are consistent with Planck
results \citep{Planck2014} with a matter density 
$\Omega_{\rm m} = 0.307$, cosmological constant
$\Omega_{\Lambda}=0.693$, dimensionless Hubble constant $h=0.678$, slope
of the power spectrum  $n=0.96$ and normalization of the power
spectrum $\sigma_{8}=0.823$.  
This translates into a particle mass of  $m_{\rm p}=1.5\times 10^{8}$
$h^{-1}$ M$_{\odot}$.

We use halo catalogs constructed with a Bound-Density-Maxima (BDM)
algorithm. 
The catalogs were obtained from the publicly available Multidark
database  \footnote{\url{http://www.multidark.org/MultiDark/}}
\citep{MultiDark}. 
For each  halo in the box we extract its comoving position and mass.  
We focus our work on halos more massive than $6.0\times
10^{9}$\hMsun\ resolved with at least  $40$ particles 
to guarantee statistical significance and a well behaved halo mass function. 
We do not take into account sub-halos.

@@

In what follows we note by the paper ${\mathcal M}$ a model
defined by a particular choice of the two parameters $M_{\rm min}$, 
$M_{\rm  max}$. For each model  ${\mathcal M}$ we define
$\tilde{f}_{\rm occ}$ as the median occupation fraction within the the
mock fields.

@@

The preferred $1-\sigma$ range for the masses is $10.0<\log \mmin<11.2$
and $11.0<\log\mmax<13.0$.
\focc\ is completely determined by \mmin\ from \focc$=$0.011 when
$\log\mmin=10.0$ to $\focc=0.39$ when $\log\mmin=11.15$. 
We compute the power-law dependence between \focc\ and \mmin\ to be 
\begin{equation}
\focc = 0.08\left(\frac{\mmin}{10^{10.3}\hMsun}\right)^{0.77}.
\end{equation}
We remind the reader that models with $\log \focc>0.00$ ($\focc>1$)
correspond to cases where the halo number density is smaller 
than the number of observed LAEs but are  still considered consistent
because of the expected uncertainty in the median number density of
LAEs due to cosmic variance (see Fig. \ref{fig:cosmicv0} and
\S\ref{subsec:explore}).

@@

@@

@@

@@

@@
